---
title: "Midterm Project"
author: "Siling Chen"
date: "October 16, 2016"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Read a data file and filter data

```{r}
rate0 <- read.csv(file.choose())
attach(rate0)
colnames(rate0) <- c("Date", "EUR", "HKD")     # change columns' names
rate <- rate0[-c(1:5),]     # remove first five rows of description
rate <- rate[rate[,2]!="ND",]     # remove NA data and create a new data set
```

## Create a new data set "eur"

```{r}
colnames(rate0) <- c("Date", "EUR", "HKD")     # change columns' names
rate <- rate0[-c(1:5),]     # remove first five rows of description
rate <- rate[rate[,2]!="ND",]     # remove NA data and create a new data set

eur <- rate[, c("Date", "EUR")]     # extract data for EURO
i <- rownames(eur)     # add an index column
eur <- cbind(i=i, eur)     # merge to a new data set
head(eur)
```

## Time series plot of EUR

Firstly, convert data from "eur" data set to time series data, and then plot. We can see the trend of daily data from 2000 to 2015 is very unstable. The currency rate starts increasing from 2001 to 2013, and then rapidly fallls. From 2004 to the end of 2015, the currency rate cycles the rapidly up-down pattern. 

```{r include=FALSE}
library(tseries)     # for time series analysis
library(dplyr)     # for manipulating data.frames
```

```{r ts_eur, echo=FALSE}
ts_eur <-  ts(eur$EUR, start = 2000, frequency = 365)     # create a time series object
plot.ts(ts_eur)     # time series plot
```


## Forecast the currency rate of EUR

After plotting the time series plot, we can use time series data to forecast the currency rate for next ten years. From the forecast plot of EUR, we can see that the trend is relatively smooth, and the currency rate would have a little decrease in the future.

```{r include=FALSE}
library(zoo)     #for handling regular and irregular time series
library(forecast)     #for forecasting functions for time series
```

```{r eur_fc, echo=FALSE}
eur_tbats = tbats(ts_eur)     # detect daily patterns in time series data
eur_fc = forecast(eur_tbats, h = 120)     # forecast EUR rate for next 120 months=10 years
plot(eur_fc)
```


## Apply ggplot2 to plot

```{r include=FALSE}
library(date)     # for handling dates
library(lubridate)     # for working with dates
library(scales)     # to access breaks/formatting functions
library(ggplot2)     # for creating graphs
library(ggthemes)     #for changing themes
```

We use boxplot to check if there is an outlier in the data set. As we see from the this plot, there is no outlier.

```{r echo=FALSE}
ggplot(eur, aes(x = 1 , y = Date)) + geom_boxplot() + xlab("Currency Rate") + ylab("Date")    # check outliers
```

Here, we use function ggplot to perform a scatterplot with x=Date and y=HKD. 

```{r echo=FALSE}
eur$Date <- as.Date(eur$Date)     #convert data class
EURplot1 <- ggplot(eur, aes(eur$Date, EUR)) + geom_point(na.rm=TRUE, color="purple", size=1) + scale_x_date(date_labels = "%b-%y") + ggtitle("EUR per USD\n 2000-2015") + xlab("Date") + ylab("Currency Rate")    # use ggplot2 to plot EUR rate
EURplot1 <- EURplot1 + theme_stata() +  stat_smooth(color = "green")    #add a theme and a trend
EURplot1
```

We can also use another plot to find the trend of EUR currency rate. In this month-year plot, we can easily see the fluctuation of the currency rate in each month of each year.

```{r echo=FALSE}
eur$year <- year(eur$Date)     # create year variables
eur$month <- month(eur$Date, label = TRUE)     # create month variables
eur <- eur[which(eur$year >= 2000), ]     # subset the data
EURplot2 <- ggplot(eur, aes(x = month, y = EUR))     # build the foundation of the plot
EURplot2 <- EURplot2 + geom_line(aes(color = factor(year), group = year)) +  scale_color_discrete(name = "Year") + labs(title = "EUR per USD\n 2000-2015", x = "Month", y = "Currency Rate")     # add lines color coded and grouped by year, name the legend "Year", format the y-axis, and add a title and labels
EURplot2
```

## Create a new data set "hkd"

```{r}
hkd <- rate[, c("Date", "HKD")]     # extract data for HKD
j <- rownames(hkd)     # add an index column
hkd <- cbind(j=j, hkd)     # merge to a new data set
head(hkd)
```

## Time series plot of HKD

At first, we also convert data from "hkd" data set to time series data, and then plot. We can see the pattern of daily data from 2000 to 2015 is much more unstable than the pattern of EUR. The currency rate stays stable from 2000 to 2002, and then rapidly falls in 2013. Until the end of 2015, the currency rate also experiences a rapidly up-town pattern. The range between the min value and the max value is very large.

```{r ts_hkd, echo=FALSE}
ts_hkd <-  ts(hkd$HKD, start = 2000, frequency = 365)     # create a time series object
plot.ts(ts_hkd)     # time series plot
```

## Forecast the currency rate of HKD

After plotting the time series plot, we can use time series data to forecast the currency rate for next ten years. From the forecast plot of HKD, we can see that the trend will be very stable, and the currency rate would almost stay in the same level in the future.

```{r hkd_fc, echo=FALSE}
hkd_tbats = tbats(ts_hkd)     # detect daily patterns in time series data
hkd_fc = forecast(hkd_tbats, h = 120)     # forecast HKD rate for next 120 months=10 years
plot(hkd_fc)
```

## Boxplot of HKD

We use a boxplot to check if there is an outlier in the data set. As we see from the plot, there is no outlier.

```{r echo=FALSE}
hkd$Date <- as.Date(hkd$Date)     #convert data class
ggplot(hkd, aes(x = 1 , y = Date)) + geom_boxplot() + xlab("Currency Rate") + ylab("Date")     # check outliers
```

## Apply ggplot2 to plot

Here, we use function ggplot to perform a scatterplot with x=Date and y=HKD. 

```{r echo=FALSE}
HKDplot1 <- ggplot(hkd, aes(hkd$Date, HKD)) + geom_point(na.rm=TRUE, color="blue", size=1) + scale_x_date(date_labels = "%b-%y") + ggtitle("HKD per USD\n 2000-2015") + xlab("Date") + ylab("Currency Rate")    # time series plot for HKD rate
HKDplot1 <- HKDplot1 + theme_stata() + stat_smooth(color = "red")     #add a theme and a trend
HKDplot1
```

We can also use another plot to find the trend of HKD currency rate. In this month-year plot, we can easily see the fluctuation of the currency rate in each month of each year.

```{r echo=FALSE}
hkd$year <- year(hkd$Date)     # create year variables
hkd$month <- month(hkd$Date, label = TRUE)     # create month variables
hkd <- hkd[which(hkd$year >= 2000), ]     # subset the data
HKDplot2 <- ggplot(hkd, aes(x = month, y = HKD))     # build the foundation of the plot
HKDplot2 <- HKDplot2 + geom_line(aes(color = factor(year), group = year)) +  scale_color_discrete(name = "Year") + labs(title = "HKD per USD\n 2000-2015", x = "Month", y = "Currency Rate")     # add lines color coded and grouped by year, name the legend "Year", format the y-axis, and add a title and labels
HKDplot2
```

## Find a relation between EUR and HKD

At first, we perform a correalation test to find if there is a correlation between EUR and HKD. Based on the Pearson's product-moment correlation test, we find that cor=-0.3298 which means EUR and HKD have a weak negative correlation.

```{r}
rate$EUR <- as.numeric(as.character(rate$EUR))     # covert data to numeric value
rate$HKD <- as.numeric(as.character(rate$HKD))
cor.test(rate$EUR, rate$HKD)     # test the correlation between EUR and HKD
```

Secondly, we build a linear regression model named m1 and perform a linear regression of EUR on HKD on the intercept. By setting the significance level ?? = 5%, we perform a statistical test on the null hypothesis that HKD is significant. From the summaries of m1, F-statistic=490.4, df=4019, and the P-value is <2e-16, which indicate HKD is significant.

```{r m1}
m1 <- lm(rate$EUR ~ rate$HKD)     # build a linear regression model
summary(m1)     #  produce summaries of the results of the model 
```

Moreover, we use ggplot2 to plot the linear regression model m1. 

```{r echo=FALSE}
rate$Date <- as.Date(rate$Date)     # cover data class
ggplot(rate, aes(rate$HKD, rate$EUR)) + geom_line() + theme_economist() +  ggtitle("Linear Regression Model \n lm(EUR ~ HKD)") + xlab("HKD") + ylab("EUR")     # use ggplot2 to plot the model
```

## Dummy variable Ti

To beign with, let Ti = i/n, i = 1,2...,4180, be a dummy variable representing time. Then, we calculate the sample correlation between EUR and the dummy variable Ti. Based on the Pearson's product-moment correlation test, we find that t-statistic=47.627 ans cor=0.6006, which states EUR and Ti have a medium positive correlation.

```{r}
dim(rate0)     # check the number of entires in each column
eur$i <- as.numeric(i)     # covert data class 
Ti = eur$i / 4180
eur <- cbind(eur, Ti)     # merge Ti to eur

eur$EUR <- as.numeric(as.character(eur$EUR))
cor.test(eur$EUR, Ti)
```

Secondly, we build a linear regression model named m2 and perform a linear regression of EUR on Ti with the intercept. By setting the significance level ?? = 5%, we perform a statistical test on the null hypothesis that Ti is not significant. From the summaries of m2, F-statistic=2268, df=4019, and the P-value is <2e-16, which indicates Ti is significant.

```{r m2}
m2 <- lm(eur$EUR ~ Ti)
summary(m2)
```

In addition, we use ggplot2 to plot the model m2.

```{r echo=FALSE}
ggplot(eur, aes(Ti, eur$EUR)) + geom_line() + theme_economist() +  ggtitle("Linear Regression Model \n lm(EUR ~ Ti)") + xlab("Time") + ylab("EUR Rate") 
```

## Dummy variable Tj

To beign with, let Tj = j/n, i = 1,2...,4180, be a dummy variable representing time. Then, we calculate the sample correlation between HKD and the dummy variable Tj. Based on the Pearson's product-moment correlation test, we find that t-statistic=-56.926 ans cor=-0.6681, which states HKD and Tj have a medium negative correlation.

```{r}
hkd$j <- as.numeric(j)     # covert data class 
Tj = hkd$j / 4180
hkd <- cbind(hkd, Tj)     # merge Tj to hkd

hkd$HKD <- as.numeric(as.character(hkd$HKD))
cor.test(hkd$HKD, Tj)
```

Secondly, we build a linear regression model named m3 and perform a linear regression of EUR on Tj with the intercept. By setting the significance level ?? = 5%, we perform a statistical test on the null hypothesis that Tj is not significant. From the summaries of m3, F-statistic=3241, df=4019, and the P-value is <2.2e-16, which indicates Tj is significant.

```{r m3}
m3 <- lm(hkd$HKD ~ Tj)
summary(m3)
```

Additionally, we use ggplot2 to plot the model m3.

```{r echo=FALSE}
ggplot(hkd, aes(Tj, hkd$HKD)) + geom_line() + theme_economist() +  ggtitle("Linear Regression Model \n lm(HKD ~ Tj)") + xlab("Time") + ylab("HKD Rate")
```

## Tidy data

Last but not least, we need to reorganize data. We merge new data to the original data set and update the final data.

```{r tidy data}
rate <- cbind(rate, Ti, Tj)
```

